\textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\ {\footnotesize Bachelorarbeit}

### 2023-05-20 Paper/chapters/abstract.tex
bibliography no ul
add abstract

Lorem ipsum dolor \cite{gift:2017} sit amet, consectetur adipisicing elit, sed do eiusmod tempor
incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu
fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
culpa qui officia deserunt mollit anim id est laborum.

>>>>

Processor designs licensed by ARM are ubiquitous in many areas due to their low
power consumption and heat generation. Being used in nearly all modern
smartphones and tablets as well as in many desktop computers and IoT devices,
the need for securing communications and confidentiality through encryption
arises. Designing cryptographic algorithms for such low-power environments is
an active area of research and various designs have been proposed. We examine
in detail the two block ciphers GIFT and Camellia. For that, we first give an
introduction of the designs as well as the ARMv8 platform we develop for. We
then present multiple implementation strategies which leverage ARMv8 SIMD
processing facilities and realize these in the C programming language using
NEON intrinsics. In the end, we present performance benchmarks and a brief
comparison to conclude this work. We find bit- and bytesliced implementations
employing vector instructions to be the fastest.

### 2023-05-20 Paper/chapters/introduction.tex
add abstract

@@

\begin{table}[h!]
    \centering
    \footnotesize
    \caption{Common NEON intrinsics}
    \begin{tabularx}{\textwidth}{llX}
        \toprule
        Intrinsic && Summary \\
        \midrule
        \texttt{uint64\_t} & \texttt{vgetq\_lane\_u64(void)} & Extract a single lane \\
        \midrule
        \texttt{void} & \texttt{vsetq\_lane\_u64(uint64\_t)} & Insert a single lane \\
        \midrule
        \texttt{uint64x2\_t} & \texttt{vdupq\_n\_u64(uint64\_t)} & Initialize all lanes to same value \\
        \midrule
        \texttt{void} & \texttt{vst1q\_u64(uint64\_t*, uint64x2\_t)} & Store from register to memory \\
        \midrule
        \texttt{uint64x2\_t} & \texttt{vld1q\_u64(uint64\_t*, uint64x2\_t)} & Load from memory to register \\
        \midrule
        \texttt{uint8x16\_t} & \texttt{veorq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise XOR \\
        \midrule
        \texttt{uint8x16\_t} & \texttt{vandq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise AND \\
        \midrule
        \texttt{uint8x16\_t} & \texttt{vorrq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise OR \\
        \midrule
        \texttt{uint8x16\_t} & \texttt{vmvnq\_u8(uint8x16\_t)} & bitwise NOT \\
        \midrule
        \texttt{uint8x16\_t} & \texttt{vqtbl2q\_u8(uint8x16\_t, uint8x16\_t)} & permutation (\texttt{TBL}) \\
        \midrule
        \texttt{uint64x2\_t} & \texttt{vextq\_u64(uint64x2\_t, uint64x2\_t, int)} & extract from pair of vectors \\
        \bottomrule
    \end{tabularx}
\end{table}

>>>>

\begin{center}
    \begin{tabular}{rl}
        $\ll_n;\gg_n$ & left; right shift by $n$ bits \\
        $\lll_n;\ggg_n$ & left; right rotate by $n$ bits \\
        $\land$ & AND \\
        $\lor$ & OR \\
        $\oplus$ & XOR \\
        $A_{(32)}$ & $A$ is 32 bits long \\
        $||$ & concatenation
    \end{tabular}
\end{center}

@@

\begin{table}[h!]
    \centering
    \scriptsize
    \caption{Common NEON intrinsics}
    \begin{tabularx}{\textwidth}{llX}
        \toprule
        Intrinsic && Summary \\
        \midrule
        \texttt{uint64x2\_t} & \texttt{vdupq\_n\_u64(uint64\_t)} & Initialize all lanes to same value \\
        \midrule
        \texttt{void} & \texttt{vst1q\_u64(uint64\_t*, uint64x2\_t)} & Store from register to memory \\
        \midrule
        \texttt{uint64x2\_t} & \texttt{vld1q\_u64(uint64\_t*, uint64x2\_t)} & Load from memory to register \\
        \midrule
        \texttt{uint8x16\_t} & \texttt{veorq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise XOR \\
        \midrule
        \texttt{uint8x16\_t} & \texttt{vandq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise AND \\
        \midrule
        \texttt{uint8x16\_t} & \texttt{vorrq\_u8(uint8x16\_t, uint8x16\_t)} & bitwise OR \\
        \midrule
        \texttt{uint8x16\_t} & \texttt{vmvnq\_u8(uint8x16\_t)} & bitwise NOT \\
        \midrule
        \texttt{uint8x16\_t} & \texttt{vqtbl2q\_u8(uint8x16\_t, uint8x16\_t)} & permutation (\texttt{TBL}) \\
        \midrule
        \texttt{uint64x2\_t} & \texttt{vextq\_u64(uint64x2\_t, uint64x2\_t, int)} & extract from pair of vectors \\
        \bottomrule
    \end{tabularx}
\end{table}
